A Spatial AR System for Wide-area Axis-aligned Metric Augmentation of Planar Scenes

### 2022-01-11 paper/elsarticle/elsarticle-template.tex
paper

\begin{figure}[t]
    \centering
    \subfloat[\centering Original image.]{\raisebox{10mm}{\includegraphics[width=5cm]{images/projim.png}}}
    \qquad
    \subfloat[\centering Projection of original image to target location in scene plane (view directly downwards to scene plane).]{{\includegraphics[width=6cm]{images/proj_projim_full_.png}}}
    \caption{Na\"ive projection to a given target location in the scene plane (e.g., the floor, a wall, or table) if not facing the scene plane directly. (a) Original image, which we wish to project to the scene plane. (b) Projecting the original image to the scene plane from the viewpoint of the projector (shown as a frustum in the bottom left, in black with up vector red) directed to the target location introduces projective distortions if the projector does not face the scene plane directly (as is the case here), with orientation and scale of the augmentation varying in the scene plane as a function of target location.}
    \label{fig:proj}
\end{figure}

@@

\begin{figure}[t]
    \centering
    \subfloat[\centering Our warped image.]{\raisebox{10mm}{\includegraphics[width=5cm]{images/projim_warped_scaled.png}}}
    \qquad
    \subfloat[\centering Projection of our warped image to target location in scene plane (view directly downwards to scene plane).]{{\includegraphics[width=6cm]{images/proj_projim_waped_full_.png}}}
    \caption{Projection to a given target location in the scene plane, correcting for projective distortions in terms of a \textit{virtual} projector placed and utilized using our approach. (a) Original image (cf.\ Figure~\ref{fig:proj}(a)) warped using our approach. (b) Projecting the warped image to the scene plane from the viewpoint of the projector (bottom left, frustum in black with up vector red; same as in Figure~\ref{fig:proj}(b)) directed to the target location has the effect of projecting the \textit{original} image from the viewpoint of a virtual projector (center, frustum in gray with up vector red). This renders the projection free of projective distortions, aligned with the axes of the virtual camera, and at the desired metric scale. Note that black pixels will in effect be ignored by the projector.}
    \label{fig:warp}
\end{figure}

@@

@@

\begin{figure}[t]
    \centering
    \subfloat[\centering Projector calibration image (2D pattern position detections overlain).]{{\includegraphics[width=5.15cm]{images/detections_.png}}}
    \quad
    \subfloat[\centering Recovered 3D positions corresponding to projector calibration image for use in projector calibration (red, in scene plane).]{{\includegraphics[height=3.7cm]{images/2d3d.png}}}
    \caption{Obtaining the 3D positions of the 2D-3D correspondences for use in projector calibration. (a) For each target location, a projector calibration image is acquired from the viewpoint of the `downwards-facing' camera, containing a projected asymmetric circle pattern at the target location and a chessboard pattern placed on the floor nearby. Circle pattern center points and chessboard corners detected automatically (overlain for illustration). (b) Given a projector calibration image, the scene plane is recovered in the coordinate frame of the camera (black with up vector red) via spatial resection with respect to 2D-3D correspondences obtained using the chessboard pattern (blue); the 3D positions of the 2D-3D correspondences to be used in calibrating the projector (red, in scene plane) are obtained by intersection with the scene plane of the back-projections (rays, likewise gray) of the 2D center points of the circle pattern detected in the image plane of the camera.} %The corresponding 2D positions are detected in the image of the asymmetric circles pattern image projected by the projector (not shown). The `downwards-facing' camera is shown relative to the scene plane in terms of its camera frustum (black, with up vector in red).
    \label{fig:2d}
\end{figure}

@@

\begin{figure}[t]
    \centering
    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=5.25cm]{images/3_proj_.pdf}}}
    \qquad
    \subfloat[\centering View directly downwards to scene~plane.]{{\includegraphics[height=5.25cm]{images/3_above3_proj.png}}}
    \caption{Obtaining the virtual projector. For a given target location, the virtual projector is obtained by (i) rotating the recovered projector (bottom left, frustum in black) about the point of intersection of its optical axis with the scene plane such that the optical axis be made parallel with the scene plane's normal vector, (ii) rotating the $X$- and $Y$-axes to align them with those of the virtual camera guaranteed---in contrast to the `downwards-facing' camera---to face \textit{directly} downwards to the scene plane (cf.\ Figure \ref{fig:virtualcam}), and (iii) translating along the normal direction to achieve the desired metric projected image dimensions. Note in (b) that the virtual projector's principal axes are aigned with those of the virtual camera in Figure~\ref{fig:virtualcam}(b).} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
    \label{fig:virtualproj}
\end{figure}

\paragraph{Virtual projector} The placement of the virtual projector---for a given a target location---determines from which pose the image to be projected is to \textit{appear} to have been projected, with the effect correcting for projective distortions, placing the axes of the augmentation in accordance with the axes of the `downwards-facing' camera, and setting the dimensions of the augmentation in line with the desired target dimensions.  We carry out this placement according to a small handful of steps. It is in accordance with the $X$- and $Y$-axes of the `downwards-facing' camera that we wish to place the $X$- and $Y$-axes of the augmentation, since we wish to intuitively be able to use the rotation of the camera relative to the scene plane to set those axes. It is, however, only if the virtual projector be placed to face the scene plane directly that the projection will appear free of distortions; this is noteworthy, since a physical camera placed to face downwards is almost certain to not face downwards to the scene plane precisely. Accordingly, we rotate the recovered camera about its center of projection to align its optical axis\footnote{Strictly speaking, the optical axis is the back-projection of the principal point; in our usage, we understand it to refer to the ray through center of the image plane.} with the normal vector of the scene plane by applying the minimum arc-length rotation that relates the two, giving a \textit{virtual} camera facing directly downwards to the scene plane (cf.\ Figure~\ref{fig:virtualcam}). To obtain the virtual projector, we then (i) intersect the scene plane with the optical axis of the projector and rotate the projector's placement about that point of intersection, aligning the optical axis with the scene plane's normal vector, and (ii) align the $X$- and $Y$-axes with those of the virtual camera. Finally, we (iii) translate along the scene plane's normal, in order to satisfy the desired metric projected image dimensions; the height $h$ [m] above the ground plane at which the virtual projector is to be placed if the image---with original image width of $w_\text{im}$ [px]---is to be projected to the scene plane with a metric width of $w$ [m] is given by
\begin{equation}
h = w \times f_\text{proj} / w_\text{im},
\end{equation}
where $f_\text{proj}$ [px] is the focal length of the projector. These three steps are illustrated in Figure~\ref{fig:virtualproj}. The resulting virtual projector is thus rendered fronto-parallel with the scene plane---enabling projection to the scene plane absent of projective distortions---and projects to the scene plane in the desired scale and with the the $X$- and $Y$-axes of the augmentation set in accordance with those of the `downwards-facing' camera. %\footnote{A physical camera placed to face downwards is almost certain to not face downwards precisely; in contrast, we align the virtual camera's optical axis exactly with the scene plane's normal vector, rendering it genuinely fronto-parallel with respect to the scene plane.}

@@

\begin{figure}
    \centering
    \subfloat[\centering Hardware setup consisting of a Panasonic PT-RZ660BE projector, a Dynamic Projection Institute steerable mirror system, and the left view of a Stereolabs Zed 2 stereo camera to serve as the `downwards-facing' camera.]{{\includegraphics[height=5cm]{images/system3_.pdf}}}  %3.6
    \quad
    \subfloat[\centering Measurement for augmentation, using a steel rule and a digital angle gauge (enlargements on the right).]{{\includegraphics[height=5cm]{images/meas2_.pdf}}}
    \caption{Evaluation scenario. (a) Our hardware setup, comprised of a projector, a steerable mirror system, and a stereo camera, of which we used only the left view. (b) Example augmentation produced by projecting---to one of the 15 target locations across the floorspace considered in our evaluation, to the upper-left corner of our floorspace---an image warped using our approach. The digital angle gauge reads $89.6^\circ{}$ for the top-left corner, and the steel rule 49.9 cm for the bottom edge. Corners of the full projected image extent projected to the floor in light green to provide context. Note that blur has a minimal impact on the quality of the augmentation, and hence on the accuracy of the measurements.}
    \label{fig:eval}
\end{figure}

>>>>

@@

\begin{figure}[t]
    \centering
    \subfloat[\centering Original image (to be projected).]{\raisebox{10mm}{\includegraphics[width=5cm]{images/projim.png}}}
    \qquad
    \subfloat[\centering Projection from side to scene plane (viewed directly downwards to scene plane).]{{\includegraphics[width=6cm]{images/proj_projim_full_.png}}}
    \caption{Na\"ive projection to a given target location in the scene plane (e.g., the floor, a wall, or table) if not facing the scene plane directly. (a) Original image, which we wish to project to the scene plane. (b) Projecting the original image to the scene plane from the viewpoint of the projector (shown as a frustum in the bottom left, in black with up vector red) directed to the target location introduces projective distortions if the projector does not face the scene plane directly (as is the case here), with orientation and scale of the augmentation varying in the scene plane as a function of target location.}
    \label{fig:proj}
\end{figure}

@@

\begin{figure}[t]
    \centering
    \subfloat[\centering Our warped image (to be projected).]{\raisebox{10mm}{\includegraphics[width=5cm]{images/projim_warped_scaled.png}}}
    \qquad
    \subfloat[\centering Projection from side to scene plane (viewed directly downwards to scene plane).]{{\includegraphics[width=6cm]{images/proj_projim_waped_full_.png}}}
    \caption{Projection to a given target location in the scene plane, correcting for projective distortions in terms of a \textit{virtual} projector placed and utilized according to our approach. (a) Original image (cf.\ Figure~\ref{fig:proj}(a)) warped using our approach. (b) Projecting the warped image to the scene plane from the viewpoint of the projector (bottom left, frustum in black with up vector red; same as in Figure~\ref{fig:proj}(b)) directed to the target location has the effect of projecting the \textit{original} image from the viewpoint of a virtual projector (center, frustum in gray with up vector red). This renders the projection free of projective distortions, aligned with the axes of the virtual camera, and at the desired metric scale. Note that black pixels will in effect be ignored by the projector.}
    \label{fig:warp}
\end{figure}

@@

\begin{figure}[t]
    \centering
    \subfloat[\centering Projector calibration image (2D pattern position detections overlain).]{{\includegraphics[width=5.15cm]{images/detections_.png}}}
    \quad
    \subfloat[\centering Recovered 3D positions corresponding to projector calibration image for use in projector calibration (red, in scene plane).]{{\includegraphics[height=3.7cm]{images/2d3d.png}}}
    \caption{Obtaining the 3D positions of the 2D-3D correspondences for use in projector calibration. (a) For each target location, a projector calibration image is acquired from the viewpoint of the `downwards-facing' camera, containing a projected asymmetric circle pattern at the target location and a chessboard pattern placed on the floor nearby. Circle pattern center points and chessboard corners detected automatically (overlain for illustration). (b) Given a projector calibration image, the scene plane is recovered in the coordinate frame of the `downwards-facing' camera (black with up vector red) via spatial resection with respect to 2D-3D correspondences obtained using the chessboard pattern (blue); the 3D positions of the 2D-3D correspondences to be used in calibrating the projector (red, in scene plane) are obtained by intersection with the scene plane of the back-projections (rays, likewise gray) of the 2D center points of the circle pattern detected in the image plane of the camera.} %The corresponding 2D positions are detected in the image of the asymmetric circles pattern image projected by the projector (not shown). The `downwards-facing' camera is shown relative to the scene plane in terms of its camera frustum (black, with up vector in red).
    \label{fig:2d}
\end{figure}

@@

\paragraph{Virtual camera and projector} The placement of the virtual projector---for a given a target location---determines from which pose the image to be projected is to \textit{appear} to have been projected, with the effect of (i) correcting for projective distortions, (ii) placing the axes of the augmentation in accordance with the axes of the `downwards-facing' camera, and (iii) setting the dimensions of the augmentation in line with the desired target dimensions.  We carry out this placement according to a small handful of steps. It is in accordance with the $X$- and $Y$-axes of the `downwards-facing' camera that we wish to place the $X$- and $Y$-axes of the augmentation, since we wish to intuitively be able to use the rotation of the camera relative to the scene plane to set those axes. It is, however, only if the virtual projector be placed to face downwards to the scene plane precisely that the projection will appear free of distortions; this is noteworthy, since a physical camera placed to face downwards is almost certain to not face precisely downwards to the scene plane. Accordingly, we rotate the recovered camera about its center of projection to align its optical axis\footnote{Strictly speaking, the optical axis is the back-projection of the principal point; in our usage, we understand it to refer to the ray through center of the image plane.} with the normal vector of the scene plane by applying the minimum arc-length rotation that relates the two, giving a \textit{virtual} camera facing directly downwards to the scene plane (cf.\ Figure~\ref{fig:virtualcam}). To obtain the virtual projector, we then (i) intersect the scene plane with the optical axis of the projector and rotate the projector's placement about that point of intersection, aligning the optical axis with the scene plane's normal vector, and (ii) align the $X$- and $Y$-axes with those of the virtual camera. Finally, we (iii) translate along the scene plane's normal, in order to satisfy the desired metric projected image dimensions; the height $h$ [m] above the ground plane at which the virtual projector is to be placed if the image---with original image width of $w_\text{im}$ [px]---is to be projected to the scene plane with a metric width of $w$ [m] is given by
\begin{equation}
h = w \times f_\text{proj} / w_\text{im},
\end{equation}
where $f_\text{proj}$ [px] is the focal length of the projector. These three steps are illustrated in Figure~\ref{fig:virtualproj}. The resulting virtual projector is thus rendered fronto-parallel with the scene plane---enabling projection to the scene plane absent of projective distortions---and projects to the scene plane in the desired scale and with the the $X$- and $Y$-axes of the augmentation set in accordance with those of the `downwards-facing' camera. %\footnote{A physical camera placed to face downwards is almost certain to not face downwards precisely; in contrast, we align the virtual camera's optical axis exactly with the scene plane's normal vector, rendering it genuinely fronto-parallel with respect to the scene plane.}

\begin{figure}[t]
    \centering
    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=5.25cm]{images/3_proj_.pdf}}}
    \qquad
    \subfloat[\centering View directly downwards to scene~plane.]{{\includegraphics[height=5.25cm]{images/3_above3_proj.png}}}
    \caption{Obtaining the virtual projector, with $X$- and $Y$-axes aligned with those of the virtual camera (cf.\ Figure~\ref{fig:virtualcam}). For a given target location, the virtual projector is obtained by (i) rotating the recovered projector (bottom left, frustum in black) about the point of intersection of its optical axis with the scene plane such that the optical axis be made parallel with the scene plane's normal vector, (ii) rotating the $X$- and $Y$-axes to align them with those of the virtual camera facing \textit{directly} downwards to the scene plane (cf.\ Figure \ref{fig:virtualcam}), and (iii) translating along the normal direction to achieve the desired metric projected image dimensions. Projection from the viewpoint of the virtual projector thus produces an augmentation absent of projective distortions, aligns the augmentation \texit{in accordance with} the horizontal and vertical image axes of the `downwards-facing' camera, and ensures the augmentation appears in the scene plane with the desired metric dimensions.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
    \label{fig:virtualproj}
\end{figure}

@@

\begin{figure}
    \centering
    \subfloat[\centering Hardware setup consisting of a Panasonic PT-RZ660BE projector, a steerable mirror system manufactured by Dynamic Projection Institute, and the left view of a Stereolabs Zed 2 stereo camera---placed to face downwards towards the scene plane---to serve as the `downwards-facing' camera.]{{\includegraphics[height=5cm]{images/system3_.pdf}}}  %3.6
    \quad
    \subfloat[\centering Measurement for augmentation, using a steel rule and a digital angle gauge (enlargements on the right).]{{\includegraphics[height=5cm]{images/meas2_.pdf}}}
    \caption{Evaluation scenario. (a) Our hardware setup, comprised of a projector, a steerable mirror system, and a stereo camera, of which we used only the left view. (b) Example augmentation produced by projecting---to one of the 15 target locations across the floorspace considered in our evaluation, to the upper-left corner of our floorspace---an image warped using our approach. The digital angle gauge reads $89.6^\circ{}$ for the top-left corner, and the steel rule 49.9 cm for the bottom edge. Corners of the full projected image extent projected to the floor in light green to provide context. Note that blur has a minimal impact on the quality of the augmentation, and hence on the accuracy of the measurements.}
    \label{fig:eval}
\end{figure}
